### 2021-04-20 doc/VerticalArrangements.tex
minor edits. Sent to Gloria and Peggy

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Data Generating Process}
This section provides an overview of our methodology, with additional details appearing in the Appendix.  We simulate markets by randomly sampling shares from a Dirichlet distribution for 2, 3, 4, or 5 retailers or wholesalers, respectively.\footnote{We parametrize the Dirichlet distribution so it is equivalent to a uniform distribution.  }  We also assume that in the pre-merger state, anywhere from 0 to 4 retailers are vertically integrated with a single wholesaler. We assume that vertically integrated wholesalers supply inputs to non-integrated retailers and that vertically integrated retailers purchase inputs from non-integrated wholesalers. The price coefficient $\alpha$ is calibrated by assuming that in the pre-merger world, there is a vertically integrated outside option available to all customers.  The other goods are differenced relative to this option, which maintains the outside good normalization.  The market size is set to 1.

@@

For each combination of number of retailers, number of wholesalers, and bargaining parameter, we draw 1,000 different sets of market primitives.  This results in 3.85 million merger simulations.  We then eliminate mergers where the merger is unprofitable to the merging firms, as well as markets that do not pass the Hypothetical Monopolist Test, yielding 2.69 million markets.\textbf{discrepancy between res.nests dimension and ALL Counts in Table 1}\footnote{The Hypothetical Monopolist Test requires that were a monopolist to jointly own all products in a candidate market, that firm would raise the price of at least one of the merging producers' products by at least a ``small but significant non-transitory increase in price'' (SSNIP), which we take to be 5\%. } All 2.69 million markets treat as primitives the number of retailers, the number of wholesalers, the bargaining parameter, and the wholesaler and retailer marginal costs, which we allow to be either constant or linear.

@@

Each panel contains four pairs of box and whisker plots, with each pair corresponding to a different type of merger. The blue box and whisker plots (on the left in each pair) depict outcomes assuming that marginal costs are constant, and the orange box and whisker plots (on the right in each pair) show outcomes assuming that marginal costs are linear.  The whiskers display the $5^{th}$ and $95^{th}$ percentiles of the outcome distribution, the boxes denote the $25^{th}$ and $75^{th}$ percentiles, and the solid horizontal line marks the median. Note that negative outcome values imply agent harm, and positive values imply agent benefits.

@@

In terms of the reltationship between bargaining power and harm, integrated mergers are perhaps most similar to vertical mergers. Like vertical mergers, there is a strong monotonic relationship between bargaining power and harm. Median consumer harm increases from about 26\% of pre-merger expenditures when retailers have relatively more bargaining power to about -1.4\% of pre-merger expenditures when wholesalers have relatively more bargaining power. Likewise, median total harm increases from 16\% of pre-merger expenditures when retailers have relatively more bargaining power to -13\% when wholesalers have relatively more bargaining power.

@@

\begin{table}
\include{sumtable}
\label{tab:simsum}
\caption{Summary Statistics}
\end{table}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Data Generating Process}
This section provides an overview of our methodology, with additional details appearing in the Appendix.  We consider upstream and downstream mergers between either two unintegrated firms or between one vertically integrated and one unintegrated firm. For vertical mergers, we only consider mergers between an unintegrated wholesaler and unintegrated retailer. Finally, we assume that an integrated merger occurs between two vertically integrated firms\textbf{consider adding diagonal mergers by allowing for heterogeneity in the bargaining parameter.}.

We simulate markets by randomly sampling shares from a Dirichlet distribution for 2, 3, 4, or 5 retailers or wholesalers, respectively.\footnote{We parameterize the Dirichlet distribution so it is equivalent to a uniform distribution.  }  Because integrated mergers must have two vertically integrated incumbents in the pre-merger state, we increase the maximum allowable number of wholesalers and retailers in our simulated markets to 7. We also assume that in the pre-merger state, there are anywhere from 0 to 4 vertically integrated incumbents (6 for integrated mergers). We assume that vertically integrated firms are no siloed: integrated wholesalers supply inputs to non-integrated retailers and that vertically integrated retailers purchase inputs from non-integrated wholesalers. The price coefficient $\alpha$ is calibrated by assuming that in the pre-merger world, there is a vertically integrated outside option available to all customers.  The other goods are differenced relative to this option, which maintains the outside good normalization.  The market size is set to 1.

@@

For each combination of number of retailers, number of wholesalers, number of incumbent integrated firms, and bargaining parameter, we draw 1,000 different sets of market primitives.  This results in 3.85 million merger simulations.  We then eliminate mergers where the merger is unprofitable to the merging firms, as well as markets that do not pass the Hypothetical Monopolist Test, yielding 2.69 million markets.\footnote{The Hypothetical Monopolist Test requires that were a monopolist to jointly own all products in a candidate market, that firm would raise the price of at least one of the merging producers' products by at least a ``small but significant non-transitory increase in price'' (SSNIP), which we take to be 5\%. } All 2.69 million markets treat as primitives the number of retailers, the number of wholesalers, the bargaining parameter, and the wholesaler and retailer marginal costs, which we allow to be either constant or linear.

@@

Each panel contains four pairs of box and whisker plots, with each pair corresponding to a different type of merger. The blue box and whisker plots (left) depict outcomes assuming that marginal costs are constant, and the orange box and whisker plots (right) show outcomes assuming that marginal costs are linear.  The whiskers display the $5^{th}$ and $95^{th}$ percentiles of the outcome distribution, the boxes denote the $25^{th}$ and $75^{th}$ percentiles, and the solid horizontal line marks the median. Note that negative outcome values imply agent harm, and positive values imply agent benefits.

@@

In terms of the relationship between bargaining power and harm, integrated mergers are perhaps most similar to vertical mergers. Like vertical mergers, there is a strong monotonic relationship between bargaining power and harm. Median consumer harm increases from about 26\% of pre-merger expenditures when retailers have relatively more bargaining power to about -1.4\% of pre-merger expenditures when wholesalers have relatively more bargaining power. Likewise, median total harm increases from 16\% of pre-merger expenditures when retailers have relatively more bargaining power to -13\% when wholesalers have relatively more bargaining power.

@@

\begin{table}
\include{sumtable}
\caption{Summary Statistics}
\label{tab:simsum}
\end{table}
